Inducing Domain Specific Languages for Bayesian Program Learning

### 2018-01-31 paper.tex
more intro

\section{Program Representation}
We choose to represent programs using $\lambda$-calculus~\cite{pierce}.
A $\lambda$-calculus expression is either:
\\\noindent A \emph{primitive}, like the number 5 or
  the function \texttt{sum}.
\\\noindent A \emph{variable}, like $x$, $y$, $z$
\\\noindent A $\lambda$\emph{-abstraction}, which creates a new function. $\lambda$-abstractions have a variable and a body. The body is a $\lambda$-calculus expression. Abstractions are written as $\lambda \text{var}. \text{body}$.
\\\noindent An \emph{application} of a function to an argument. Both the function and the argument are $\lambda$-calculus expressions. The application of the function $f$ to the argument $x$ is written as $f\; x$.

@@

\begin{algorithm}[tb]
   \caption{Grammar Induction Algorithm}
   \label{grammarInductionAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
     \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$, AIC coefficient $a$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
     \STATE Define $\log \probability[\mathcal{D}]\stackrel{+}{ = } -\lambda\sum_{p\in \mathcal{D}} \text{size}(p)$
     \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
     \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - a|\mathcal{D}|$
     \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
     \WHILE {true}
     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ a subtree of }z\}$
     \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
     \IF{$\text{score}(\mathcal{D}') > \text{score}(\mathcal{D})$}
     \STATE $\mathcal{D}\gets\mathcal{D}'$
     \ELSE
     \STATE\textbf{return} $\mathcal{D},\theta^*(\mathcal{D})$
     \ENDIF
     \ENDWHILE
   \end{algorithmic}
\end{algorithm}

>>>>

\subsection{Motivation and Overview}

\system takes as input a set of \emph{tasks}, written $X$, each of which is a program induction problem.
It has at its disposal a \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.
Its goal is to solve each of the tasks by writing a program,
and also to infer a DSL $\mathcal{D}$ that distills the commonalities across all of the programs that solve the tasks.

We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Fig.~\ref{graphicalModel}. From Fig.~\ref{graphicalModel}, the MAP probability of $(\mathcal{D},\theta)$ is
\begin{equation}
  \probability[\mathcal{D},\theta|X]\propto \probability[\mathcal{D},\theta] \prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\label{probabilityDSL}
\end{equation}
If we had a $(\mathcal{D},\theta)$ maximizing Eq.~\ref{probabilityDSL}, then we could recover the most likely program for task $x$ by maximizing $\probability[x|p] \probability[p|\mathcal{D},\theta]$.
Through this lens we now take as our goal to maximize Eq.~\ref{probabilityDSL}.
But even \emph{evaluating} Eq.~\ref{probabilityDSL} is intractable because it involves summing over the infinite set of all possible programs. In general, programs are hard-won: finding even a single program that explains a given observation presents a daunting combinatorial search problem.

With this fact in mind, we will instead maximize the following tractable lower bound on Eq.~\ref{probabilityDSL},
which we call $J$:
\begin{equation}
J = \probability[\mathcal{D},\theta]\prod_{x\in X}\log \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]\label{lowerBound}
\end{equation}
This lower bound depends on sets of programs, $\left\{\mathcal{F}_x \right\}_{x\in X}$:

\noindent\textbf{Definition.} The \emph{frontier of task $x$}, written $\mathcal{F}_x$,
is the set of programs discovered by \system where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.

We maximize $J$ by alternatingly maximizing it w.r.t. the DSL and the frontiers:
\\\noindent \textbf{Program Synthesis: Maximizing $J$ w.r.t. the frontiers.} Here we
want to find new programs to add to  the frontiers so that $J$ increases the most.
Adding new programs to the frontiers means attempting to synthesize new programs for each of the tasks.
So interleaved with the DSL induction steps are program induction steps. Section~\ref{explorationSection}
explains how \system synthesizes new programs and given a DSL.
\\\noindent \textbf{DSL Induction: Maximizing $J$ w.r.t. the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed and so we can evaluate $J$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $J$.  Section~\ref{grammarInductionSection} explains this step of the algorithm.

In practice, synthesizing programs is extremely difficult because
of how large the search space is. Learning the DSL eases the difficulty of synthesis
by exposing a domain-specific basis for constructing programs.
Another complementary means of meeting search is to learn a
bottom-up \emph{recognition model}

\subsection{Synthesizing Programs from a DSL}\label{explorationSection}

Describe the enumerator, depth first search on the random choices made by the generative model with iterative deepening

Limitation: has a hard time learning arbitrary constants. Solution: solve those constants using specialized solvers.

\subsection{Learning a Neural Recognition Model}
The key idea here is that if the DSL is suitably fit to the domain then
we can get away with a simple, low-capacity model.

\subsection{Inducing the DSL from the Frontiers}\label{grammarInductionSection}

\begin{algorithm}[tb]
   \caption{Grammar Induction Algorithm}
   \label{grammarInductionAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
     \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$, AIC coefficient $a$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
     \STATE Define $\log \probability[\mathcal{D}]\stackrel{+}{ = } -\lambda\sum_{p\in \mathcal{D}} \text{size}(p)$
     \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
     \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - a|\mathcal{D}|$
     \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
     \WHILE {true}
     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ a subtree of }z\}$
     \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
     \IF{$\text{score}(\mathcal{D}') > \text{score}(\mathcal{D})$}
     \STATE $\mathcal{D}\gets\mathcal{D}'$
     \ELSE
     \STATE\textbf{return} $\mathcal{D},\theta^*(\mathcal{D})$
     \ENDIF
     \ENDWHILE
   \end{algorithmic}
\end{algorithm}
\section{Program Representation}
We choose to represent programs using $\lambda$-calculus~\cite{pierce}.
A $\lambda$-calculus expression is either:
\\\noindent A \emph{primitive}, like the number 5 or
  the function \texttt{sum}.
\\\noindent A \emph{variable}, like $x$, $y$, $z$
\\\noindent A $\lambda$\emph{-abstraction}, which creates a new function. $\lambda$-abstractions have a variable and a body. The body is a $\lambda$-calculus expression. Abstractions are written as $\lambda \text{var}. \text{body}$.
\\\noindent An \emph{application} of a function to an argument. Both the function and the argument are $\lambda$-calculus expressions. The application of the function $f$ to the argument $x$ is written as $f\; x$.

@@
